---
title: "GPDI517: Introduction to data.table"
author: "by Tugba Ozturk"
date: "April 4, 2019"
output:
  prettydoc::html_pretty:
    theme: cayman
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **data.table** is another R library to analyze data sets!

Let's start with installing and loading the `data.table` library:

```{r installation}
#install.packages("data.table") #installs the R package
library("data.table") #loads the package
```

## It turns data frames into data tables to perform faster manipulation.

To create a local data table for our practices, we can use the `airquality` data frame:

```{r create-mydf, echo=TRUE}
mydf <- data.table(airquality)
```

Note that the `echo` parameter can be set to `FALSE` to prevent printing of the R code in the pdf document. While practicing, we will keep the code chunks visible. 

To print the data table, use the following: 

```{r print-mydf}
print(mydf)
```

A data table is an enhanced data frame. When printed, there is no need to use `head()` function.

Let's subset `mydf` to see which values (rows) have the `Wind` variable larger than 8 in September:

```{r subsetting}
mydf[Wind >8 &  Month==9L]
```

The following lines would sort `mydf` first by `Day` column and then by `Temp`:

```{r sorting}
mydf[order(Day,-Temp)]
```

Note that `Day` and `Temp` are in ascending and descending orders, respectively since `Temp` is given with a `-` sign.

Here, add comment lines to explain what the following code chunk does:

```{r practice1}
mydf[,Temp]
mydf[,c("Day","Temp")]
```

Here, add comment lines explaining how the following lines of code **differ** from one another:

```{r practice2}
mean(mydf[,Ozone])
mydf[,mean(Ozone, na.rm=TRUE)]
mydf[mean(Ozone, na.rm=TRUE)]
mydf[,.(mean(Ozone, na.rm=TRUE))]
mydf[,.(mean(Ozone, na.rm=TRUE)), by=Month]
```


Run the following code and fix it if necessary:

```{r practice3}
mydf[,.(aveOz = mean(Ozone), sd(Ozone)), by=Month]
```
